Loop-Aware Optimizations in PyPy's Tracing JIT

### 2013-01-09 talk/dls2012/dls04-ardo.tex
py3k's December update

\begin{abstract}
One of the nice properties of a tracing JIT is that many of its optimizations
are simple, requiring one forward pass only. This is not true for loop-invariant code
motion which is a very important optimization for code with tight kernels.
Especially for dynamic languages that typically perform quite a lot of loop invariant
type checking, boxed value unwrapping and virtual method lookups.

@@

One of the advantages that tracing JIT compilers have above traditional
method-based
JITs is that their optimizers are much easier to write. Because a tracing JIT
produces only linear pieces of code without control flow joins, many
optimization passes on traces can have a very simple structure: They often
consist of one forward pass replacing operations by faster ones or even
discarding them as they walk along it. This makes
optimization of traces very similar to symbolic execution. Also, many
difficult problems in traditional optimizers become tractable if the optimizer
does not need to deal with control flow merges.

@@

The work described in this paper was done in the context of the PyPy
project.\footnote{\texttt{http://pypy.org}} PyPy is a framework for implementing
dynamic languages efficiently~\cite{rigo_pypys_2006}. When implementing a
language with PyPy, one writes an interpreter for the language in RPython~\cite{ancona_rpython:_2007}.
RPython (``Restricted Python``) is a subset
of Python chosen in such a way that it can be efficiently translated to a
C-based VM by performing type inference.

>>>>

\begin{abstract}
One of the nice properties of a tracing just-in-time compiler (JIT) is that many of its optimizations
are simple, requiring one forward pass only. This is not true for loop-invariant code
motion which is a very important optimization for code with tight kernels.
Especially for dynamic languages that typically perform quite a lot of loop invariant
type checking, boxed value unwrapping and virtual method lookups.

@@

One of the advantages that tracing just-in-time compilers (JITs) have above traditional
method-based
JITs is that their optimizers are much easier to write. Because a tracing JIT
produces only linear pieces of code without control flow joins, many
optimization passes on traces can have a very simple structure: They often
consist of one forward pass replacing operations by faster ones or even
discarding them as they walk along it. This makes
optimization of traces very similar to symbolic execution. Also, many
difficult problems in traditional optimizers become tractable if the optimizer
does not need to deal with control flow merges.

@@

The work described in this paper was done in the context of the PyPy
project.\footnote{\texttt{http://pypy.org}} PyPy is a framework for implementing
dynamic languages efficiently~\cite{rigo_pypys_2006}. When implementing a
language with PyPy, one writes an interpreter for the language in RPython~\cite{ancona_rpython:_2007}.
RPython (``Restricted Python``) is a subset
of Python chosen in such a way that it can be efficiently translated to a
C-based virtual machine (VM) by performing type inference.
